---
output: 
  pdf_document: 
    latex_engine: xelatex
---

<!--
You can delete the header-includes (lines 3-5 above) if you like and also the chunk below since it is loaded in the skeleton.Rmd file.  They are included so that chap2.Rmd will compile by itself when you hit Knit PDF.
-->

```{r include_reedtemplates_2, include = FALSE}
# This chunk ensures that the reedtemplates package is installed and loaded
# This reedtemplates package includes the template files for the thesis and also
# two functions used for labeling and referencing
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")

if(!require(reedtemplates)){
  library(devtools)
  devtools::install_github("ismayc/reedtemplates")
  }
library(reedtemplates)
```


# The Adventures of Mark Twain {#math-sci}

Now that I've gone through \textit{The Adventures of Tom Sawyer}, I wanted to take a look at Twain's works as a whole (using the six novels stated in **Chapter 1**) and compare them to important milestones in his life.  I'll use the same methods as I did in **Chapter 1**, and in **Chapter 3** I will introduce a different method for textual analysis.  

### Important dates in Twain's life

Twain led a very exciting and inventive life.  When he was young, his family moved to the city of Hannibal, MO.  Hannibal was a frequent stop for steam boats, and it came to be quite the inspiration for Twain later in his works.  Twain was educated at a local private school until he droppped out to become a printer's apprentice, where he discovered his love for writing.  Later in his life, while living in California, Twain met his future wife, Olivia Langdon.  They had one son and three daughters together, only of which one of them would make it to adulthood.  Twain frequently invested in new inventions, and towards the end of his career, he was broke from poor investments.  Below is a timeline of Twain's life in picture format.  

```{r, fig.width=6, fig.height=3,echo=FALSE}
library(png)
library(grid)
img <- readPNG("twaintimeline.png")
grid.raster(img)
```

### Twain's life as seen through his works

#### Sentiment Analysis

```{r, include = FALSE}
library(mosaic) 
library(XML)
library(readr)
library(devtools)
library(stringr)
library(RCurl)
library(syuzhet)
library(tm)
```

```{r, include = FALSE}
# importing texts
roughingitURL <- "http://www.gutenberg.org/files/3177/3177.txt"
roughingittext <- scan(roughingitURL, what = "character", sep = "\n")
innocentsURL <- "http://www.gutenberg.org/files/3176/3176-0.txt"
innocentstext <- scan(innocentsURL, what = "character", sep = "\n")
huckfinnURL <- "http://www.gutenberg.org/files/76/76-0.txt"
huckfinntext <- scan(huckfinnURL, what = "character", sep = "\n")
tomsawyerURL <- "http://www.gutenberg.org/files/74/74-0.txt"
tomsawyertext <- scan(tomsawyerURL, what = "character", sep = "\n")
connyankeeURL <- "http://www.gutenberg.org/files/86/86-0.txt"
connyankeetext <- scan(connyankeeURL, what = "character", sep = "\n")
puddURL <- "http://www.gutenberg.org/files/102/102-0.txt"
puddtext <- scan(puddURL, what = "character", sep = "\n")
```

```{r, include = FALSE}
# cleaning roughing it text
findstart <- function(text){ 
  startend <- 0
  for (i in 1:length(text)){
    if (text[i] == "CHAPTER I.") {
      startend <- i
    }
  }
  for (j in 50:length(text)){
    if (text[j] == "End of Project Gutenberg’s Roughing It, by Mark Twain (Samuel Clemens)") {
      startend <- c(startend, j)
      return(startend)
    }
  }
} 

actualtext <- function(text, startend){
  realtext <- text[startend[1]:startend[2]]
  return (realtext)
}

# roughing it text
roughstartend <- findstart(roughingittext)
roughingittext <- actualtext(roughingittext, roughstartend)
```

```{r, include=FALSE}
# cleaning innocents abroad text
findstart <- function(text){ 
  startend <- 0
  for (i in 1:length(text)){
    if (text[i] == "CHAPTER I.") {
      startend <- i
    }
  }
  for (j in 50:length(text)){
    if (text[j] == "End of the Project Gutenberg EBook of The Innocents Abroad") {
      startend <- c(startend, j)
      return(startend)
    }
  }
} 

actualtext <- function(text, startend){
  realtext <- text[startend[1]:startend[2]]
  return (realtext)
}

# innocents abroad text
innocentstartend <- findstart(innocentstext)
innocentstext <- actualtext(innocentstext, innocentstartend)
```

```{r, include=FALSE}
# cleaning tom sawyer text
findstart <- function(text){ 
  startend <- 0
  for (i in 1:length(text)){
    if (text[i] == "CHAPTER I") {
      startend <- i
    }
  }
  for (j in 50:length(text)){
    if (text[j] == "CONCLUSION") {
      startend <- c(startend, j)
      return(startend)
    }
  }
} 

actualtext <- function(text, startend){
  realtext <- text[startend[1]:startend[2]]
  return (realtext)
}

# tom sawyer text
tomsawyerstartend <- findstart(tomsawyertext)
tomsawyertext <- actualtext(tomsawyertext, tomsawyerstartend)
tomsawyertext <- iconv(tomsawyertext,"WINDOWS-1252","UTF-8")
```

```{r, include = FALSE}
# cleaning huckleberry finn text
findstart <- function(text){ 
  startend <- 0
  for (i in 1:length(text)){
    if (text[i] == "CHAPTER I.") {
      startend <- i
    }
  }
  for (j in 50:length(text)){
    if (text[j] == "End of the Project Gutenberg EBook of Adventures of Huckleberry Finn,") {
      startend <- c(startend, j)
      return(startend)
    }
  }
} 

actualtext <- function(text, startend){
  realtext <- text[startend[1]:startend[2]]
  return (realtext)
}

# huckleberry finn text
huckfinnstartend <- findstart(huckfinntext)
huckfinntext <- actualtext(huckfinntext, huckfinnstartend)
```

```{r, include = FALSE}
# cleaning connecticut yankee text
findstart <- function(text){ 
  startend <- 0
  for (i in 1:length(text)){
    if (text[i] == "CHAPTER I") {
      startend <- i
    }
  }
  for (j in 50:length(text)){
    if (text[j] == "End of the Project Gutenberg EBook of A Connecticut Yankee in King Arthur’s") {
      startend <- c(startend, j)
      return(startend)
    }
  }
} 

actualtext <- function(text, startend){
  realtext <- text[startend[1]:startend[2]]
  return (realtext)
}

# connecticut yankee text
connyankeestartend <- findstart(connyankeetext)
connyankeetext <- actualtext(connyankeetext, connyankeestartend)
```

```{r, include=FALSE}
# cleaning pudd'n text
findstart <- function(text){ 
  startend <- 0
  for (i in 1:length(text)){
    if (text[i] == "CHAPTER I.") {
      startend <- i
    }
  }
  for (j in 50:length(text)){
    if (text[j] == "and the creditors sold him down the river.") {
      startend <- c(startend, j)
      return(startend)
    }
  }
} 

actualtext <- function(text, startend){
  realtext <- text[startend[1]:startend[2]]
  return (realtext)
}

# pudd'n text
puddstartend <- findstart(puddtext)
puddtext <- actualtext(puddtext, puddstartend)
```

```{r, include = FALSE}
# sentiment across entire life
early <- c(roughingittext, innocentstext)
earlylen <- length(early)
middle <- c(huckfinntext, tomsawyertext)
midlen <- length(middle)
late <- c(connyankeetext, puddtext)

allworks <- c(early, middle, late)

allworkssent <- get_sentiment(allworks, "syuzhet")
allworkssenttrans <- as.numeric(get_transformed_values(allworkssent, 
                                  low_pass_size = 3,
                                  scale_vals = TRUE,
                                  scale_range = FALSE))
allworksdata <- data.frame(cbind(linenumber = seq_along(allworkssenttrans), ft = allworkssenttrans))
```

```{r, fig.align="center", cache=TRUE, warning=FALSE, echo = FALSE}
ggplot(data = allworksdata, aes(x = linenumber, y = ft)) +
        geom_bar(stat = "identity", alpha = 0.8, color = "midnightblue", fill = "midnightblue") +
        theme_minimal() +
        ylab("Transformed Sentiment Value") +
        ggtitle(expression(paste("Sentiment in Twain's works "))) +
        theme(axis.title.x=element_blank()) +
        theme(axis.ticks.x=element_blank()) +
        theme(axis.text.x=element_blank()) +
  ggplot2::annotate("text", size = 3, x = c(20, 50, 80, 35, 55, 77, 90), 
                 y = c(-2, -2, -2, -1, 1, 1.2, -1),
                 label = c("Early works", "Middle works", "Late works", "Marriage", "Death of son", "Founded publishing\n company", "Bankruptcy")) + 
  ggplot2::annotate("segment", arrow = arrow(length = unit(0.03, "npc")),x = c(35, 55, 77, 90), xend = c(35, 55, 77, 90),y = c(-.9, .9, 1, -.9),yend = c(-.1, .1, .2, -.1))
```

As expected, milestones in Twain's life match up with the sentiment throughout his works.  Appropriately, the sentiment reaches an all time high around the time he meets his wife, Olivia, and gets married.  The sentiment plummets after the death of their first child and never quite recovers to the original happiness.  The works end on a sad note when half his children have died and he's bankrupt.  

#### Stylometric Analysis

```{r, include = FALSE}
library(katherinemansfieldr)
# sentence length
sentlen <- function(sentences) {
  senlen <- 0
  for (i in 1:length(sentences)) {
    sentlen <- sapply(gregexpr("\\W+", sentences[i]), length) + 1
    senlen <- c(senlen, sentlen)
  }
  senlen <- senlen[-1]
  return(senlen)
}

earlySent <- extract_sentences(early)
earlysenlen <- sentlen(earlySent)
earlysendata <- as.data.frame(cbind(earlysenlen, c(1:length(earlysenlen))))

middleSent <- extract_sentences(middle)
middlesenlen <- sentlen(middleSent)
middlesendata <- as.data.frame(cbind(middlesenlen, c(1:length(middlesenlen))))

lateSent <- extract_sentences(late)
latesenlen <- sentlen(lateSent)
latesendata <- as.data.frame(cbind(latesenlen, c(1:length(latesenlen))))
```

```{r, include = FALSE}
library(ggplot2)
# early sentence length
earlysenlenplot <- ggplot(earlysendata, 
                    aes(x = V2, y = earlysenlen)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(earlysendata$earlysenlen), 
             colour = "red", linetype = "dotted") +
  ylab("Average Sentence Length \n (# of words)") + 
  xlab("Line Number") +
  ggtitle("Average Sentence Length \n for Early Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include=FALSE}
# middle sentence length
middlesenlenplot <- ggplot(middlesendata, 
                    aes(x = V2, y = middlesenlen)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(middlesendata$middlesenlen), 
             colour = "red", linetype = "dotted") +
  ylab("Average Sentence Length \n (# of words)") + 
  xlab("Line Number") +
  ggtitle("Average Sentence Length \n for Middle Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include = FALSE}
# late sentence length
latesenlenplot <- ggplot(latesendata, 
                    aes(x = V2, y = latesenlen)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(latesendata$latesenlen), 
             colour = "red", linetype = "dotted") +
  ylab("Average Sentence Length \n (# of words)") + 
  xlab("Line Number") +
  ggtitle("Average Sentence Length \n for Late Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include=FALSE}
# token length
tokenlen <- function(token) {
  toklen <- 0
  for (i in 1:length(token)) {
    tokenlen <- nchar(token[i])
    toklen <- c(toklen, tokenlen)
  }
  toklen <- toklen[-1]
  return(toklen)
}

earlyToken <- extract_token(early)
earlytokenlength <- tokenlen(earlyToken)
earlytokendata <- as.data.frame(cbind(earlytokenlength, c(1:length(earlytokenlength))))

middleToken <- extract_token(middle)
middletokenlength <- tokenlen(middleToken)
middletokendata <- as.data.frame(cbind(middletokenlength, c(1:length(middletokenlength))))

lateToken <- extract_token(late)
latetokenlength <- tokenlen(lateToken)
latetokendata <- as.data.frame(cbind(latetokenlength, c(1:length(latetokenlength))))
```

```{r, include = FALSE}
# early token length
earlytokenplot <- ggplot(earlytokendata, 
                    aes(x = V2, y = earlytokenlength)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(earlytokendata$earlytokenlength), 
             colour = "red", linetype = "dotted") +
  ylab("Average Token Length \n (# of characters)") + 
  xlab("Line Number") +
  ggtitle("Average Token Length \n for Early Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include = FALSE}
# middle token length
middletokenplot <- ggplot(middletokendata, 
                    aes(x = V2, y = middletokenlength)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(middletokendata$middletokenlength), 
             colour = "red", linetype = "dotted") +
  ylab("Average Token Length \n (# of characters)") + 
  xlab("Line Number") +
  ggtitle("Average Token Length \n for Middle Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include= FALSE}
# late token length
latetokenplot <- ggplot(latetokendata, 
                    aes(x = V2, y = latetokenlength)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(latetokendata$latetokenlength), 
             colour = "red", linetype = "dotted") +
  ylab("Average Token Length \n (# of characters)") + 
  xlab("Line Number") +
  ggtitle("Average Token Length \n for Late Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include = FALSE}
# early quotation
earlypunc <- freq_punct_line(early, "“")

earlyquoteplot <- ggplot(earlypunc, 
                    aes(x = line_index, y = left_quote)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(earlypunc$left_quote), 
             colour = "red", linetype = "dotted") +
  ylab("Average Number of\n Speaking Lines") + 
  xlab("Line Number") +
  ggtitle("Average Number of Speaking\n Lines for Early Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include = FALSE}
# middle quotation
middlepunc <- freq_punct_line(middle, "“")

middlequoteplot <- ggplot(middlepunc, 
                    aes(x = line_index, y = left_quote)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(middlepunc$left_quote), 
             colour = "red", linetype = "dotted") +
  ylab("Average Number of\n Speaking Lines") + 
  xlab("Line Number") +
  ggtitle("Average Number of Speaking\n Lines for Middle Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

```{r, include = FALSE}
# late quotations
latepunc <- freq_punct_line(late, "“")

latequoteplot <- ggplot(latepunc, 
                    aes(x = line_index, y = left_quote)) + 
  geom_smooth() + 
  geom_hline(yintercept = mean(latepunc$left_quote), 
             colour = "red", linetype = "dotted") +
  ylab("Average Number of\n Speaking Lines") + 
  xlab("Line Number") +
  ggtitle("Average Number of Speaking\n Lines for Late Stories") + 
  theme(axis.title.y = element_text(size = 9)) + 
  theme(axis.title.x = element_text(size = 9)) + 
  theme(plot.title = element_text(size = 12))
```

As we can see from the plots below, general trends between early, middle, and late works do not seem to differ drastically.  One difference to note among sentence length is that Twain's earlier works seem to have longer sentences in general, which isn't surprising because his earlier works described landscapes and nature scenery.  His later works had a larger variety in sentence length, which could be due to becoming a more critical and experienced writer over the years.  
Among token length, the general trend across is that tokens are longer as the stories progress.  The earlier works again have longer tokens than middle or late works, and that could be due to the emphasis on nature and scenery in those works.
There are more speaking lines in the middle works than early or late works.  Again this isn't surprising because of the content in those works.  The trends are different between the three phases, speaking lines increase as the novels progress in early and late works, but decrease for the middle works.  
Because there were only two works used for each period, this analysis is of course limited because I didn't analyze Twain's full body of work for each period.  However, I did pick his most notable pieces to analyze, so these graphs could be able to say something about the general essence of his writing from those periods.  


```{r, echo = FALSE, warning=FALSE}
library(cowplot)
library(ggplot2)
plot_grid(earlysenlenplot, middlesenlenplot, latesenlenplot, 
          earlytokenplot, middletokenplot, latetokenplot,
          earlyquoteplot, middlequoteplot, latequoteplot,
          nrow = 3, ncol = 3)
```





